---
title: "test"
author: "alexM"
date: "2026-04-28"
output: html_document
---

# loading and connecting database
```{r}
# Load libraries and database
library(DBI)
library(RSQLite)
library(tidyverse)
 
db_path <- "C:/Users/Sanik/AppData/Local/activitywatch/activitywatch/aw-server/peewee-sqlite.v2.db"
con <- dbConnect(RSQLite::SQLite(), dbname = db_path) # connect to db
# the connect object must be passed to all other functions to do with the db
```

# exploring the database
```{r}
print(dbListTables(con)) # list all contained tables
print(dbListFields(con,"eventmodel")) # list the fields, "columns", in a given table
df = dbReadTable(con,"eventmodel") # read entire table
apps = dbGetQuery(con, "SELECT * FROM eventmodel WHERE bucket_id = 1") # this is an SQL query
websites = dbGetQuery(con, "SELECT timestamp, duration, datastr FROM eventmodel WHERE bucket_id = 3")
dbDisconnect(con)
```
# cleaning and saving rds
```{r}
#lets theoretically clean the data:
# Convert timestamp to datetime
websites$timestamp <- as.POSIXct(websites$timestamp, format = "%Y-%m-%d %H:%M:%S")
# Convert duration to numeric if needed
websites$duration <- as.numeric(websites$duration)
#something like that
#now we can save the cleaned data as an rds:
saveRDS(websites, "data/websites_test.rds")
# and we can read that like this:
# df = readRDS("data/websites_test.rds")
```
```{r}
library(tidyverse)
library(lubridate)
apps = readRDS("data/apps_clean.rds")
websites = readRDS("data/websites_clean.rds")
#lets separate the column event_start:

apps <- apps %>% 
  mutate(week=floor_date(event_start, unit="week"),
         day = floor_date(event_start, unit="day"),
         month = floor_date(event_start, unit="month")) %>% 
  mutate(category = case_when(is.na(category) ~ "Other", 
                              category == "browser" ~ "Browser",
                                TRUE ~ category))

websites <- websites %>% 
  mutate(week=floor_date(event_start, unit="week"),
         day = floor_date(event_start, unit="day"),
         month = floor_date(event_start, unit="month")) %>% 
  mutate(category = case_when(is.na(category) ~ "Other", TRUE ~ category))

total_time_apps = apps %>% 
  group_by(week, category) %>%
  summarise(total_time=sum(active_duration))

#now a stacked colourful barchart for the categories:
total_time_apps %>% 
  #filter(category=="Browser") %>% 
  ggplot(aes(x= week, y= total_time/3600, fill=category)) +
  theme_minimal()+
  geom_bar(stat="identity")+
  labs(y="Total time (hours)", x="Week")

```


```{r}
total_time_apps %>% group_by(category) %>% 
  summarise(sum(total_time))
```


```{r}
apps = readRDS("data/apps_clean.rds")

dotw_counts = apps %>%
  mutate(day  = floor_date(event_start, "day"),
         dotw = lubridate::wday(event_start, label = TRUE)) %>%
  distinct(day, dotw) %>%
  count(dotw, name = "n_days")

test = apps %>%
  mutate(
    day  = floor_date(event_start, "day"),
    dotw = lubridate::wday(event_start, label = TRUE)
  ) %>%
  group_by(day, hour, dotw) %>%
  summarise(hour_total = sum(active_duration), .groups = "drop") %>%
  # NO per-day normalization — keep raw durations
  group_by(hour, dotw) %>%
  summarise(total = sum(hour_total), .groups = "drop") %>%
  # Divide by number of that weekday in the data
  left_join(dotw_counts, by = "dotw") %>%
  mutate(mean_duration = total / n_days) %>%
  complete(hour = 0:23, dotw, fill = list(mean_duration = 0))

test %>%
  ggplot(aes(x = hour, y = dotw, height = mean_duration, fill = dotw)) +
  geom_density_ridges(stat = "identity", scale = 2.5, alpha = 0.8) +
  scale_x_continuous(breaks = seq(0, 23, 3),
                     labels = \(x) sprintf("%02d:00", x)) +
  scale_fill_viridis_d(option = "magma", guide = "none") +
  labs(x = NULL, y = NULL,
       title = "App usage across the day",
       subtitle = "Mean duration per hour (seconds)") +
  theme_classic()

library(plotly)

p = test %>%
  ggplot(aes(x = hour, y = dotw, height = mean_duration, fill = dotw)) +
  geom_density_ridges(stat = "identity", scale = 3, alpha = 0.7) +
  scale_x_continuous(breaks = seq(0, 23, 3),
                     labels = \(x) sprintf("%02d:00", x)) +
  scale_fill_viridis_d(option = "magma", guide = "none") +
  labs(x = NULL, y = NULL,
       title = "App usage across the day",
       subtitle = "Mean duration per hour") +
  theme_classic()

ggplotly(p)
```


# Business Hours Productivity Analysis

Analyzing laptop usage during business hours only (weekdays, 9 AM - 5 PM) provides a clearer picture than all-time statistics, which conflate necessary downtime with actual work distractions.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)

apps = readRDS("data/apps_clean.rds")

websites = readRDS("data/websites_clean.rds")

productive_categories = c("Learning & Education", "Development & Programming", "Productivity & Work", "Art & Creative")
productive_web_categories = c("Communication", "Reference & Research", "Productivity & Work", "Development & Programming", "Learning & Education", "Job Search", "AI Tools", "Science & Academia", "Government & Administration")

apps = apps %>% 
  mutate(productive = case_when(category %in% productive_categories ~ TRUE, TRUE ~ FALSE)) %>% 
  filter(category != "Browser") %>% #since browser is already counted in websites
  select(-app)

websites = websites %>% 
  mutate(productive = case_when(category %in% productive_web_categories ~ TRUE, TRUE ~ FALSE)) %>% 
  select(-domain)

# Filter to business hours: weekdays before 5 PM
business_hours = apps %>% 
  bind_rows(websites) %>%
  mutate(
    weekday_num = lubridate::wday(event_start, week_start=1)
  ) %>%
  filter(
    hour < 17,
    weekday_num >= 1 & weekday_num <= 5
  ) %>% 
  filter(hour > 6)

# Weekly aggregation
weekly_business = business_hours %>%
  group_by(week = floor_date(event_start, "week", week_start=1), productive) %>%
  summarise(total_time = sum(active_duration) / 3600, .groups = "drop") %>%
  pivot_wider(names_from = productive, values_from = total_time, values_fill = 0) %>%
  rename(unproductive = `FALSE`, productive = `TRUE`) %>%
  mutate(
    total = productive + unproductive,
    productivity_ratio = round(productive / total * 100, 1)
  )

weekly_business %>%
  ggplot(aes(x = week)) +
  geom_area(aes(y = productive, fill = "Productive"), alpha = 0.75) +
  geom_area(aes(y = unproductive, fill = "Unproductive"), alpha = 0.75) +
  geom_line(aes(y = productive + unproductive), color = "black", size = 0.3, alpha = 0.5) +
  scale_fill_manual(values = c("Productive" = "#2ecc71", "Unproductive" = "#e74c3c")) +
  labs(
    title = "Business Hours Laptop Usage (Weekdays 9 AM - 5 PM)",
    subtitle = "14-week period (Feb - May 2026)",
    x = "Week Number",
    y = "Hours",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", margin = margin(b = 5)),
    plot.subtitle = element_text(size = 11, color = "gray40", margin = margin(b = 10)),
    legend.position = "bottom",
    legend.direction = "horizontal",
    panel.grid.major.x = element_blank()
  )
```

```{r, warning=FALSE, message=FALSE}
# Summary stats
business_hours %>%
  group_by(productive) %>%
  summarise(
    total_hours = sum(active_duration) / 3600,
    .groups = "drop"
  )
```


## Key Metrics
Important to note - I excluded stuff like "New Tab" in the browser, because i dont think thats productive or unproductive, but it was actually a significant part of the browser time apparently, like 20 hours over 3 months. Perhaps this is an issue with how ActivityWatch records data.

During business hours, **56.7% of laptop time is productive**, with the following breakdown:

Also important to note is that from April 1st, My part-time research assisstant contract ended, which impacted the productive time greatly.

- **129.5 hours productive** across 14 weeks (~9.25 hours/week)
- **98.5 hours unproductive** across 14 weeks (~7.04 hours/week)  

## What Steals Your Focus

```{r, warning=FALSE, message=FALSE}
category_business = business_hours %>%
  group_by(category, productive) %>%
  summarise(hours = sum(active_duration) / 3600, sessions = n(), .groups = "drop") %>%
  arrange(desc(hours)) %>% 
  mutate(category = case_when(is.na(category)~"Misc (Domains below top 250)",
                                TRUE ~ category))

top_drains_business = category_business %>%
  filter(productive == FALSE) %>%
  arrange(desc(hours)) %>%
  head(8)

top_drains_business %>%
  ggplot(aes(x = reorder(category, hours), y = hours, fill = category)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Business Hours Productivity Drains",
    subtitle = "Unproductive activities during work hours (9 AM - 5 PM, weekdays)",
    x = NULL,
    y = "Hours"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 13, face = "bold", margin = margin(b = 5)),
    plot.subtitle = element_text(size = 10, color = "gray40", margin = margin(b = 10)),
    axis.text.y = element_text(size = 10)
  )
```

**The leverage point:** Entertainment & Leisure (25.7 hours) and Gaming (21.0 hours) account for **46.7 hours (39%)**(that needs to be recalculated) of unproductive work time. Eliminating these would boost your productivity ratio from 56.7% to approximately RECALCULATE%.

Entertainment is the #1 drain during business hours, followed by gaming.

## Where Productive Time Goes

```{r, warning=FALSE, message=FALSE}
productive_business = category_business %>%
  filter(productive == TRUE) %>%
  arrange(desc(hours)) %>%
  head(8)

productive_business %>%
  ggplot(aes(x = reorder(category, hours), y = hours, fill = category)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Productive Activities During Business Hours",
    subtitle = "What you spend work-hour time on when focused",
    x = NULL,
    y = "Hours"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 13, face = "bold", margin = margin(b = 5)),
    plot.subtitle = element_text(size = 10, color = "gray40", margin = margin(b = 10)),
    axis.text.y = element_text(size = 10)
  )
```

Productive work is concentrated in:
- **Productivity & Work tools** (primary focus) - mainly made up of google docs - that was me writing scientific articles while I had a Research assisstant job
- **Development & Programming** (strong presence)
- **AI Tools** used for development and programming

## Insights & Takeaways

1. **Business hours performance is reasonable:** 49.7% productivity is a respectable baseline, not a crisis. RECALCULATE

2. **Volatility is the real pattern:** Weeks range from 13.7% to 70.2%—the gap between bad and good weeks is huge. High-productivity weeks (60%+) are clearly achievable. RECALCULATE and 

3. **Entertainment is the immediate target:** Many entertainment hours. WHile gaming has been largely eliminated, an alarming rise in social media and entertainment (mainly youtube.com) is driving an unproductivity surge.

```{r}
df <- weekly_business |>
  arrange(week) |>
  mutate(tooltip = paste0(
    "<b>Week of ", format(week, "%b %d"), "</b><br>",
    "Productive: ", round(productive, 1), "h<br>",
    "Unproductive: ", round(unproductive, 1), "h<br>",
    "Total: ", round(total, 1), "h  •  ", productivity_ratio, "% productive"
  )) |>
  pivot_longer(c(productive, unproductive), names_to = "type", values_to = "hours") |>
  mutate(type = factor(str_to_title(type), levels = c("Unproductive", "Productive")))

glimpse(df)
head(df)

ggplot(df, aes(x = week, y = hours, fill = type)) +
  geom_area(position = "identity", alpha = 0.6)+
  scale_fill_manual(values = c("Productive" = "#2ecc71", "Unproductive" = "#e74c3c")) +
  scale_x_datetime(date_labels = "%b %d") +
  #chart_theme() +
  labs(x = NULL, y = "Hours")

ggplotly(p, tooltip = "text") |>
  apply_dark_layout() |>
  config(displayModeBar = FALSE)

```
```{r}
apps = readRDS("data/apps_clean.rds")
test = apps |> 
  mutate(week=floor_date(event_start, "week",week_start=1))|>
  group_by(week,category) |>
  summarise(total=sum(active_duration)/3600, .groups="drop") |>
  complete(week, category, fill = list(total = 0)) |>
  filter(category == "Gaming") |>
  slice_head(n = 4) |>
  pull(total)


avg_last4 <- apps |>
      mutate(week = floor_date(event_start, "week", week_start=1)) |>
      group_by(week, category) |>
      summarise(total = sum(active_duration) / 3600, .groups = "drop") |>
      complete(week, category, fill = list(total = 0)) |>
      filter(category == "Gaming") |>
      slice_tail(n = 4) |>
      pull(total) |> mean()
```

